Literature Review

### 2019-02-25 Text_classifier_models.Rmd
update

```{r, child="_setup.Rmd"}
```

#Overview
This page provides general background on theory and approaches to text classification. Links below connect to sub-pages that describe more specific topics. Follow the links for details.

Topic Categorization and Modeling Methods: https://adanieljohnson.github.io/default_website/Topic_categorization.html

Choosing a Text Classifier Method: https://adanieljohnson.github.io/default_website/Choosing_classifier.html

Designing a Classifier Evaluation Workflow: https://adanieljohnson.github.io/default_website/Classifier_comparison.html

Strategies for Improving, Evaluating Text Classifiers: https://adanieljohnson.github.io/default_website/Improving_classifier.html

@@

#Basic Workflow
Text classification problems can be broken down into four overlapping sub-problems. Each piece may be more or less important to the overall problem, but all need to be considered:

1.  Feature selection and engineering
2.  Topic categorization and modeling
3.  Selecting an appropriate text classifier method
4.  Evaluating the outcomes

@@

##Feature Selection and Engineering
This is the process of selecting, tabulating, calculating, or creating text-based features for subsequent evaluation. This includes:

*  Identifying existing, directly observable text features within the dataset
*  Engineering additional features using established NLP methods
*	 Reverse engineering relevant features based on an empirically defined feature set.

The features listed below are used regularly for natural language processing. Not all will be appropriate for all situations though; selecting, evaluating, comparing, and validating the appropriate features for the classification task is an iterative process, not a one-time decision.

*  Count-based Vectors
	+  Word Count – total number of words in the documents
	+  Character Count – total number of characters in the documents
	+  Punctuation Count – total number of punctuation marks in the documents
	+  Upper Case Count – total number of upper case words in the documents
	+  Title Word Count – total number of consecutive upper case (title) words in the documents
*  Distribution and Frequency Calculations
	+  Average word density of sentences 
	+  Average length of the words used in the documents
	+  TF-IDF
	+  Reading level estimates (described below)
*  Part of Speech Tagging and Frequency Distribution of Tags 
	+ Noun Count
	+ Verb Count
	+ Adjective Count
	+ Adverb Count
	+ Pronoun Count
*  Word Embedding (described below)

Additional features are available for longer texts such as multi-sentence paragraphs.

*  Word choices
*  Lexical diversity (described below)
*  Stance patterns - hedging, booster, generalization words or phrasing 
*  Concision - comparisons of noun phrase density versus comment clause density
*  Cohesion - use of words that are additive (also), countering (however), specify or emphasize (even, both, especially), connecting (former, latter)
*  Civility and uncertainty - language that acknowledges limits of statement, critiques ideas not people, and makes fewer generalizations

@@

###Lexical Diversity
This is not the same as word choices. This explanation is an excerpt from "quanteda: Quantitative Analysis of Textual Data", which describes how to use ```textstat_lexdiv``` to calculate lexical diversity. 
https://rdrr.io/cran/quanteda/man/textstat_lexdiv.html
http://quanteda.io/reference/textstat_lexdiv.html

@@

koRpus package in R also provides functions to estimate lexical diversity or complexity.
https://subscription.packtpub.com/book/big_data_and_business_intelligence/9781783551811/2/ch02lvl1sec16/lexical-diversity

@@

###Readability metrics
These estimate the relative grade level or age of a typical reader who could comprehend 80+% (typically) of the text. Most readability metrics require longer blocks of text (100+ words) to achieve a stable score. They use some combination of word count, average syllables per word, and sentence length to estimate reading level.

@@

###Word Embedding
**Word embedding** is a set of language modeling and feature learning techniques in natural language processing (NLP) where words or phrases from the vocabulary are mapped to vectors of real numbers. Conceptually it involves a mathematical embedding from a space with one dimension per word to a continuous vector space with a much lower dimension.

@@

One of the main limitations of word embeddings (word vector space models in general) is that possible meanings of a word are conflated into a single representation (a single vector in the semantic space). Sense embeddings are a solution to this problem: individual meanings of words are represented as distinct vectors in the space.

>>>>

```{r, child="_setup.Rmd"}
```
This page provides general background on theory and approaches to text classification. Links below connect to sub-pages that describe more specific topics. Follow the links for details.

@@

#Basic Workflow
Text classification problems can be broken down into four overlapping sub-problems. Each piece may be more or less important to the overall problem, but all need to be considered.

1.  Feature selection and engineering (described here)
2.  Topic categorization and modeling (https://adanieljohnson.github.io/default_website/Topic_categorization.html)
3.  Selecting an appropriate text classifier method (https://adanieljohnson.github.io/default_website/Choosing_classifier.html)
4.  Evaluating the outcomes, and improving the classifier ( https://adanieljohnson.github.io/default_website/Improving_classifier.html)

@@

#Feature Selection and Engineering
This is the process of selecting, tabulating, calculating, or creating text-based features for subsequent evaluation. This includes:

*  Identifying existing, directly observable text features within the dataset
*  Engineering additional features using established NLP methods
*	 Reverse engineering relevant features based on an empirically defined feature set

The features listed below are used routinely in natural language processing. Not all will be appropriate for all situations though; **selecting, evaluating, comparing, and validating appropriate features for a classification task is an iterative process, not a one-time decision.**

*  Count-based Vectors
	+  Word Count – total number of words in the documents
	+  Character Count – total number of characters in the documents
	+  Punctuation Count – total number of punctuation marks in the documents
	+  Upper Case Count – total number of upper case words in the documents
	+  Title Word Count – total number of consecutive upper case (title) words in the documents
*  Distribution and Frequency Calculations
	+  Average word density of sentences 
	+  Average length of the words used in the documents
	+  TF-IDF
	+  Reading level estimates, readability (described below)
*  Part of Speech Tagging and Frequency Distribution of Tags 
	+ Noun Count
	+ Verb Count
	+ Adjective Count
	+ Adverb Count
	+ Pronoun Count
*  Word Embedding (described below)

Additional features are available for longer texts such as multi-sentence paragraphs. These features require analysis of sentence structure, syntax, parts of speech patterns, etc.

*  Word choices (vocabulary range)
*  Lexical diversity (described below)
*  Stance patterns - hedging, booster, generalization words or phrasing 
*  Concision - comparisons of noun phrase density versus comment clause density
*  Cohesion - use of words that are additive (also), countering (however), specify or emphasize (even, both, especially), connecting (former, latter)
*  Civility and uncertainty - language that acknowledges limits of statement, critiques ideas not people, and makes fewer generalizations

@@

#Some Feature Engineering Methods
##Lexical Diversity
This is different from simple word frequencies The following explanation is an excerpt from "quanteda: Quantitative Analysis of Textual Data", which describes how to use ```textstat_lexdiv``` to calculate lexical diversity. 
https://rdrr.io/cran/quanteda/man/textstat_lexdiv.html
http://quanteda.io/reference/textstat_lexdiv.html

@@

koRpus package in R also provides functions to estimate lexical diversity or complexity. For more about this package, go to 
https://subscription.packtpub.com/book/big_data_and_business_intelligence/9781783551811/2/ch02lvl1sec16/lexical-diversity

@@

##Readability metrics
These estimate the relative grade level or age of a typical reader who could comprehend 80+% (typically) of the text. Most readability metrics require longer blocks of text (100+ words) to achieve a stable score. They use some combination of word count, average syllables per word, and sentence length to estimate reading level.

@@

##Word Embedding
**Word embedding** is a set of language modeling and feature learning techniques in natural language processing (NLP) where words or phrases from the vocabulary are mapped to vectors of real numbers. The vectors then serve as engineered features for subsequent classifier functions.

Conceptually it involves a mathematical embedding from a space with one dimension per word to a continuous vector space with a much lower dimension.

@@

One of the main limitations of word embeddings (and word vector space models in general) is that possible meanings of a word are conflated into a single representation (a single vector in the semantic space). Sense embeddings are a solution to this problem: individual meanings of words are represented as distinct vectors in the space.

### 2019-02-25 build.Rmd
updates
update

---
title: "The Build"
author: "Dan Johnson"
date: "2/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Overview
This page outlines the process for building various components used in the project. Details of each  component are provided in these sub-pages.

Structure of the initial dataset: https://adanieljohnson.github.io/default_website/datastructure.html

Import and post-import data validation: https://adanieljohnson.github.io/default_website/dataimport.html

>>>>

---
title: "The Build"
date: "2/22/2019"
bibliography: library.bib
biblio-style: apalike
---

```{r, child="_setup.Rmd"}
```

This page links to sub-pages describing how various components used in the overall project were constructed. Follow the links for details.

Extracting TAs' Comments (***Draft***): https://adanieljohnson.github.io/default_website/docxcommentextraction.html

Codebook for Classifying TAs' Comments: https://adanieljohnson.github.io/default_website/codebook.html

Structure of Development Dataset; Data Import & Validation: https://adanieljohnson.github.io/default_website/dataimport.html

Initial Exploration of the Dataset: https://adanieljohnson.github.io/default_website/initial_exploration.html

Current Version of the Naive Bayes Comment Classifier: https://adanieljohnson.github.io/default_website/NB_2r2c.html

Embedding the Naive Bayes Classifier in an Ensemble Model (***Draft***): https://adanieljohnson.github.io/default_website/Ensemble_Notes.html

### 2019-02-25 initial_exploration.Rmd
update

#Overview
The first question to answer before making deeper comparisons is, **are word frequencies different between TAs' comments from different sub-categories?** This page describes the methods used for this initial exploration.

Results of these analyses are discussed on this page: https://adanieljohnson.github.io/default_website/findings.html

@@

#The Analyses
I ran three different analyses of word usage frequency.

*  __Pearson correlations for word use frequencies__. This only compares words found in all of the Subject sub-categories
*  __Word frequency differences based on log-odds ratios__. This highlights terms that are the most different between sub-categories.
*  __N-grams mapping__ (*draft*). This identifies text strings, phrases, and terms that are unique (or nearly unique) to one sub-category.

@@

#ALTERNATIVE METHOD for pulling out stopwords.  
sortedwords.comments.all <- base_data_tokenized %>%
  filter(!word %in% stop_words$word,
         !word %in% str_remove_all(stop_words$word, "'"),
         str_detect(word, "[a-z]")) %>%
  count(word, word, sort = TRUE) %>%
  ungroup()

@@

The next block joins the individual data tables for each subset in a larger file using column named 'word'. Columns get confusing names so they are renamed. 
```{r}
total <- merge(sortedwords.comments.all.fraction,sortedwords.technical.fraction,by="word") 
total2 <- merge(total,sortedwords.writing.fraction,by="word")
total3 <- merge(total2,sortedwords.logic.fraction,by="word")
#rename lines that were possibly confusing
names(total3)[2] <- "all.count"
names(total3)[4] <- "technical.count"
names(total3)[6] <- "writing.count"
names(total3)[8] <- "logic.count"

@@

The following six graphs compare frequencies for ~770 words that appear in ALL three sub-categories: Writing, Technical, and Logic.
```{r}
ggplot(total6, aes(`all.fraction`,`writing.fraction`)) +
  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  geom_abline(color = "red")

@@

#Log-Odds Ratios of Word Frequencies
##Which Words Are Shared LEAST Between Categories?
Unlike the preceding analysis, log-odds ratio comparison does not eliminate unshared terms. Instead, it identifies which words are MOST likely to be used in one sub-category, but not another. For simplicity I reduced the analysis to just the four most common comment subjects. This has been the most informative preliminary visualization I have found so far.

This code block is based on a method for comparing word usage frequencies between two separate Twitter users [https://www.tidytextmining.com/twitter.html](described here.) If the words used by TAs in their comments differ between sub-categories, there should be fairly low Pearson correlations between word frequencies.

>>>>

The first question to answer before making deeper comparisons is, **are word frequencies different between TAs' comments from different sub-categories?** This page describes the methods used for initial exploration of the TA comments dataset.

Results of these analyses are discussed in depth on this page: https://adanieljohnson.github.io/default_website/findings.html

@@

#The Analyses
I ran three different analyses of word use frequency. For simplicity I reduced the analysis to just the four most common comment subjects.

*  __Pearson correlations for word use frequencies__. This method compares the **absolute frequencies** for words found in **ALL** of the Subject sub-categories.
*  __Word frequency differences based on log-odds ratios__. This method highlights terms that are the **most different between sub-categories**. Again, this method compares frequencies for words found in **ALL** of the Subject sub-categories.
*  __N-grams mapping__ (***Draft***). This method identifies text strings, phrases, and terms that are unique (or nearly unique) to one sub-category. Unlike the other two methods, this one does not drop terms which are missing from one or more groups.

@@

#Stopwords removal.  
sortedwords.comments.all <- base_data_tokenized %>%
  filter(!word %in% stop_words$word,
         !word %in% str_remove_all(stop_words$word, "'"),
         str_detect(word, "[a-z]")) %>%
  count(word, word, sort = TRUE) %>%
  ungroup()

@@

The next block joins the individual data tables for each subset in a larger file using column named 'word'. Columns get confusing names so they are renamed. The graphs compare frequencies for ~770 words that appear in ALL three sub-categories: Writing, Technical, and Logic.

```{r}
total <- merge(sortedwords.comments.all.fraction,sortedwords.technical.fraction,by="word") 
total2 <- merge(total,sortedwords.writing.fraction,by="word")
total3 <- merge(total2,sortedwords.logic.fraction,by="word")
#rename lines that were possibly confusing
names(total3)[2] <- "all.count"
names(total3)[4] <- "technical.count"
names(total3)[6] <- "writing.count"
names(total3)[8] <- "logic.count"

@@

#Plot graphs
ggplot(total6, aes(`all.fraction`,`writing.fraction`)) +
  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  geom_abline(color = "red")

@@

#Log-Odds Ratios of Word Frequencies
##Which Words Are Shared LEAST Between Categories?
Log-odds ratio comparison identifies the words that have the greatest difference in usage between the two groups being compared. This code block is based on a method for comparing word usage frequencies between two separate Twitter users [https://www.tidytextmining.com/twitter.html](described here.) If the words used by TAs in their comments differ between sub-categories, there should be fairly low Pearson correlations between word frequencies.

### 2019-02-25 litreview.Rmd
correction
update

#Overview
The supporting literature for this project is broad and convoluted. Rather than try to compress it to a single page, I have provided links out to sub-pages that describe various background topics. Follow the links for details.

@@

Primer on Text Classification Strategies:  https://adanieljohnson.github.io/default_website/Text_classifier_models.html

>>>>

#Overview
This project brings together several traditionally separate fields of research:

*  General learning theories
*  Educational research specific to STEM fields
*  Writing in the disciplines, and technical writing instruction
*  Text analysis and classification

Instead of trying to weave a coherent one-page narrative of these very different fields, I have provided links out to sub-pages that describe various background topics separately. Follow the links for details.

General Learning Theories Relevant to the Current Challenge:
https://adanieljohnson.github.io/default_website/pagename.html

Educational Research Specific to STEM fields
https://adanieljohnson.github.io/default_website/pagename.html

@@

Primer on Text Classification Strategies; Feature Selection and Engineering:  https://adanieljohnson.github.io/default_website/Text_Classifier_Models.html

>Sub-pages:  
>1.  Topic categorization and modeling (https://adanieljohnson.github.io/default_website/Topic_categorization.html)  
>2.  Selecting an appropriate text classifier method (https://adanieljohnson.github.io/default_website/Choosing_classifier.html)  
>3.  Evaluating the outcomes, and improving the classifier (https://adanieljohnson.github.io/default_website/Improving_classifier.html)
